\Title

### 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
captions

We also observed that the distribution of the kernel's weights is sparse, with most values close to zero (\seeFig{kernels}). As shown in the formalization of our event-based model, the computational cost of our model, if implemented on a neuromorphic chip, would be dominated by the computations used for the convolution operation. In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel. But if the information support is sparse, then computations can be now performed only on those events. Specifically, if we set some weights of the kernels to zero, then the additive operation in the convolution for those addresses can be dropped. As a consequence, computations will be performed only on those events which were multiplexed by the pruned connectivity matrix. Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights. This hypothesis is consistent with biological observations which have shown that communication consumes 35 times more energy than computation in the human cortex~\citep{Levy2021}.

@@

    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line). The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel. If we shorten the length of the kernel by using only the weights at the shortest delays, the accuracy quickly drops. However, if we prune the lowest coefficients from the whole kernel, we observe a stable accuracy value, with a drop to half-saturation observed at about $375$ times fewer computations.
    \label{fig:accuracy}
By selectively setting certain weights to zero, we observe that the accuracy's evolution, as a function of the logarithmic percentage of active weights, aligns well with a sigmoid curve for both pruning and shortening strategies. The shortening strategy (depicted in orange) demonstrates a rapid decline in accuracy, reaching half-saturation when approximately one-third of the weights are removed. On the other hand, the pruning strategy (shown in blue) exhibits a different behavior. It reaches half-saturation when the ratio of active weights is approximately $2.6\times 10^{-3}$, corresponding to around $375$ less computations compared to the dense scenario. In comparison to using the complete kernels, our method maintains accuracy close to its peak performance even when the number of computations is divided by a factor of up to around 31. This substantial reduction in computations showcases the robustness of the presented method.
\subsection{Testing with natural-like textures}
%: fig:motion_clouds
    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
    % TODO : show prototypes of textures on top
    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from a grating (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to grating (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)). Examples snapshots are shown as an illustration in the top insets. Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) and in the generic case where the orientation is independent of direction ('aperture').
    \label{fig:motion_clouds}
In order to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model, we now test the model on simpler, parameterized stimuli. For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control relevant parameters for motion detection, including motion direction, spatial orientation, and spatial frequency along with their respective precisions (\seeFig{motion_clouds})~\citep{Leon2012,vacher_bayesian_2018}. By matching the spatial and temporal characteristics of the generated movies with those of the motion task mentioned earlier, we created a range of textures featuring different spatial properties and motions. This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $12$ directions and $3$ speeds. For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability. This method provides a rich dataset of textured movies for which we know the ground truth for the motion.

We observe some interesting facts. First, as we change the mean spatial frequency of the texture, we observe a broadly tuned response in accuracy. This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006,Perrinet2007} and reveals the most informative scales learned by our model. Then, by modifying the bandwidth in spatial frequency, we show that the accuracy is worse for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}. Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth. When the stimulus is grating-like and the orientation is arbitrary with respect to the direction of motion, the system faces the aperture problem and experiences a sharp decrease in accuracy. This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion. Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter. This shows a progressive decrease in accuracy, similar to that observed in the amplitude of human eye movements~\citep{mansour_pour_speed_2018}.
%
\section{Discussion}
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic heterogeneous delay SNN that was trained using supervised learning on a visual motion detection task. We have shown how the model performs well compared to other event-based classification algorithms for this task. The learned model has a number of similarities with neurobiological and behavioral observations. The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining robust classification performance. This shows that we can exploit the precise timing of spikes to improve the performance of neural computations.
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
The HD-SNN model was trained and evaluated on a naturalistic motion detection task with realistic eye movements. It is defined such as to provide an optimal detection of  spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}. We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware. We show that the use of heterogeneous delays may be an efficient computational solution for future neuromorphic hardware, but also a key to understanding why spikes are a universal component of neural information processing.

>>>>

\begin{figure}%[h!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
    \caption{
        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line). The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel. If we shorten the length of the kernel by using only the weights at the shortest delays, the accuracy quickly drops. However, if we prune the lowest coefficients from the whole kernel, we observe a stable accuracy value, with a drop to half-saturation observed at about $375$ times fewer computations.
        }
    \label{fig:accuracy}
\end{figure}
%
We also observed that the distribution of the kernel's weights is sparse, with most values close to zero (\seeFig{kernels}). As shown in the formalization of our event-based model, the computational cost of our model, if implemented on a neuromorphic chip, would be dominated by the computations used for the convolution operation. In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel. But if the information support is sparse, then computations can be now performed only on those events. Specifically, if we set some weights of the kernels to zero, then the additive operation in the convolution for those addresses can be dropped. As a consequence, computations will be performed only on those events which were multiplexed by the pruned connectivity matrix. Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights. This hypothesis is consistent with biological observations which have shown that communication consumes 35 times more energy than computation in the human cortex~\citep{Levy2021}.

@@

By selectively setting certain weights to zero, we observe that the accuracy's evolution, as a function of the logarithmic percentage of active weights, aligns well with a sigmoid curve for both pruning and shortening strategies. The shortening strategy (depicted in orange) demonstrates a rapid decline in accuracy, reaching half-saturation when approximately one-third of the weights are removed. On the other hand, the pruning strategy (shown in blue) exhibits a different behavior. It reaches half-saturation when the ratio of active weights is approximately $2.6\times 10^{-3}$, corresponding to around $375$ less computations compared to the dense scenario. In comparison to using the complete kernels, our method maintains accuracy close to its peak performance even when the number of computations is divided by a factor of up to around 31. This substantial reduction in computations showcases the robustness of the presented method.
%
%
%%%-----------------------------------------------------------------
%
\subsection{Testing with natural-like textures}
%%%-----------------------------------------------------------------
In order to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model, we now test the model on simpler, parameterized stimuli. For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control relevant parameters for motion detection, including motion direction, spatial orientation, and spatial frequency along with their respective precisions (\seeFig{motion_clouds})~\citep{Leon2012,vacher_bayesian_2018}. By matching the spatial and temporal characteristics of the generated movies with those of the motion task mentioned earlier, we created a range of textures featuring different spatial properties and motions. This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $12$ directions and $3$ speeds. For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability. This method provides a rich dataset of textured movies for which we know the ground truth for the motion.

%%%-----------------------------------------------------------------
%: fig:motion_clouds
%%%-----------------------------------------------------------------
\begin{figure}%[h!]
    \centering
    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
    % TODO : show prototypes of textures on top
    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from gratings (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to gratings (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)). Examples snapshots are shown as an illustration in the top insets. Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) and in the generic case where the orientation is independent of direction ('aperture').
    }
    \label{fig:motion_clouds}
\end{figure}
%%%-----------------------------------------------------------------
We observe some interesting facts. First, as we change the mean spatial frequency of the texture, we observe a broadly tuned response in accuracy. This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006,Perrinet2007} and reveals the most informative scales learned by our model. Then, by modifying the bandwidth in spatial frequency, we show that the accuracy is worse for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}. Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth. When the stimulus is grating-like and the orientation is arbitrary with respect to the direction of motion, the system faces the aperture problem and experiences a sharp decrease in accuracy. This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion. Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter. This shows a progressive decrease in accuracy, similar to that observed in the amplitude of human eye movements~\citep{mansour_pour_speed_2018}.
%
\section{Discussion}
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic heterogeneous delay SNN that was trained using supervised learning on a visual motion detection task. We have shown how the model performs well compared to other event-based classification algorithms for this task. The learned model has a number of similarities with neurobiological and behavioral observations. The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining robust classification performance. This shows that we can exploit the precise timing of spikes to improve the performance of neural computations.
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
The HD-SNN model was trained and evaluated on a naturalistic motion detection task with realistic eye movements. It is defined such as to provide an optimal detection of  spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}. We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware. We show that the use of heterogeneous delays may be an efficient computational solution for future neuromorphic hardware, but also a key to understanding why spikes are a universal component of neural information processing.
